---
title: "Your title"
author: "Your name goes here"
date: "2025-11-15"
mainfont: Times New Roman
format:
    pdf:
      documentclass: article
      geometry:
        - margin=1in
      fontsize: 12pt
      linestretch: 2
      include-in-header:
        text: |
          \usepackage{etoolbox}
          \AtBeginEnvironment{verbatim}{\linespread{1}\selectfont\small}
          \AtBeginEnvironment{Highlighting}{\linespread{1}\selectfont\small}
      include-before-body: 
        text: |
          \begin{center}
          \text{SDS6/12345/2025}
          
          \vspace{0.5em}
          
          \text{Department of Mathematics}
          
          \text{Faculty of Science and Technology}
          \end{center}
          \vspace{2em}
editor: visual
bibliography: example_references.bib
link-citations: true
---

\newpage
\tableofcontents
\newpage

# Introduction

Originally developed in the early 1990s by Ross Ihaka and Robert Gentleman to support the teaching of introductory statistics, R [@rcoreteam2025b] has since evolved into a key language in data science, widely used for data cleaning, analysis, visualization, and reporting of results...

It also compares the approaches recommended in the book with my prior practices for instance, contrasting the *tidyverse* philosophy with the *data.table* approach, to illustrate how these perspectives align or differ in the context of data **manipulation** and *analysis*.

# Overview of the chapters

## Chapter 1: Data Visualization

The book opens with an introduction to data visualization in R. To use the package *ggplot2* in R, you can install it directly or install it through *tidyverse*. *tidyverse* is a collection of packages in R that are used for importing, cleaning, modelling and visualizing data in R. You can install packages as follows:

```{r}
installed.packages("tidyverse")
```

Once the package is installed, you must "call" it in R to use the functions that it provides. We do that using the function `library()` as follows;

```{r}
library(tidyverse)
```

Tidyverse produces some messages and warning once you load it, and in Quarto, you can silence those messages using the `#| warning: false` code chunk option as follows:

```{r}
#| warning: false
library(tidyverse)
```

There a lot more options for Quarto code chunks and you can learn more about them [using this link](https://quarto.org/docs/computations/execution-options.html#figure-options)

using the *ggplot2* ... mostly using the function `ggplot() +`. Here, we practice with the data `mtcars` in R [^1]

[^1]: The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

```{r}
#| label: fig-scatter-plot
#| fig-cap: "A scatter plot of Miles Per Gallon aganist Weight "
#| 
library(ggplot2)
ggplot(mtcars, aes(x = mpg, y = wt)) +
  geom_point() +
  theme_minimal()
```

The above code chunk produced @fig-scatter-plot. You car read more about [Quarto from this link](https://quarto.org).

\newpage

# References
